---
title: "Overall metagenome statistics"
author: "Sam Barnett"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
    toc: true
    toc_depth: 2
---

# Introduction

Included here is the inital analysis of the metagenome data. We are specifically examining broad trends from reads such as estimated genome size, diversity, and taxonomic makeup.

## Librarys and global variables

Here are some libraries used in this analysis and the global varaibles that will be used throughout. Mostly variables for consistent plotting.
```{r, warning=FALSE, message=FALSE}
# Libraries for data
library(dplyr)
library(phyloseq)
library(ape)
library(readxl)

# Libraries for analysis
library(vegan)
library(picante)
library(Nonpareil)
library(lme4)
library(lmerTest)
              
# Libraries for plotting
library(ggplot2)
source("/Users/sambarnett/Documents/Misc_code/paul_tol_colors.R")

# Functon for extracting legends
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)} 


# Site lists
used_sites = c("Cen08", "Cen11", "Cen14", "Cen15", "Cen16", "Cen17", "Cen19", 
               "Cen21", "Cen22", "Cen23")

# Setting repeated plot aesthetics
## Sites
site.col = paultol_colors(length(used_sites))
names(site.col) = used_sites

site.shape = c(22, 24, 24, 24, 24, 22, 24, 24, 24, 22)
names(site.shape) = used_sites

## Fire Classifications
FC.col = c("FireAffected" = "red", "Reference" = "grey")
FC.shape = c("FireAffected" = 24, "Reference" = 22)

# Basic plotting theme so as not to continually repeat it
publication_theme = theme_bw() +
  theme(axis.text = element_text(size=6),
        axis.title = element_text(size=7),
        legend.text = element_text(size=6),
        legend.title = element_text(size=7, hjust=0.5),
        strip.text = element_text(size=7),
        plot.title = element_text(size=8, hjust=0.5))

present_theme = theme_bw() +
  theme(axis.text = element_text(size=10),
        axis.title = element_text(size=12),
        legend.text = element_text(size=10),
        legend.title = element_text(size=12, hjust=0.5),
        strip.text = element_text(size=10),
        plot.title = element_text(size=14, hjust=0.5))


```

## Metadata

Read in the metadata.
```{r, warning=FALSE, message=FALSE}
# Sample metadata
sample.meta = read_xlsx("/Users/sambarnett/Documents/Shade_lab/Centralia_project/Centralia_soil_metadata.xlsx", 
                        sheet = "Metagenomic_samples", na="NA") %>%
  filter(SampleID != "Cen08_07102019_R1") %>%
  arrange(SiteID, Year) %>%
  mutate(Seq_number = row_number()) %>%
  mutate(nonpareil_file = paste("/Users/sambarnett/Documents/Shade_lab/Centralia_project/Metagenomics/Data/nonpareil/", SampleID, "_S", Seq_number, ".npo", sep=""),
         SequenceID = paste(SampleID, Seq_number, sep="_S"))

readstats.df = read.table("/Users/sambarnett/Documents/Shade_lab/Centralia_project/Metagenomics/Data/Filtered_read_statistics.txt", 
                          header=TRUE, sep="\t", comment.char = "", quote = "") %>%
  mutate(total_reads = Forward_read_count + Reverse_read_count) %>%
  select(SampleID, total_reads) %>%
  rename(SequenceID = SampleID) 

mapped_reads.df = read.table("/Users/sambarnett/Documents/Shade_lab/Centralia_project/Metagenomics/Data/Mapped_read_totals.txt", 
                          header=TRUE, sep="\t", comment.char = "", quote = "")
```


# Metagenome coverage and diversity

Lets see how well we sequenced these metagenomes. We used nonpareil to estimate the coverage of these metagenomes as well as estimate diversity.

## Get nonpareil data
Load the data. For the coverages at different sequencing depths we need to convert coverage curves into a dataframe format we can play around in ggplot with.
```{r, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=7}
# Read in the the nonpareil data. These are separate files.
nonpareil.full = Nonpareil.set(sample.meta$nonpareil_file)

# Get the coverage curves
nonpareil.df = data.frame()
for (i in seq(1, length(nonpareil.full@np.curves))){
  nonpareil.df = rbind(nonpareil.df, 
                       data.frame(SequenceID = nonpareil.full@np.curves[[i]]$label,
                                  effort = nonpareil.full@np.curves[[i]]$x.adj,
                                  coverage = nonpareil.full@np.curves[[i]]$y.cov))
}

# Add in the metadata
nonpareil.df = nonpareil.df %>%
  left_join(sample.meta, by = "SequenceID")

# Get the summary table that includes coverage (at full sequencing depth) and diversity, along with other metrics.
nonpareil.sum = data.frame(summary(nonpareil.full)) %>%
  tibble::rownames_to_column(var="SequenceID") %>%
  mutate(SampleID = gsub("_S.*", "", SequenceID)) %>%
  left_join(sample.meta, by = c("SequenceID", "SampleID"))
```

## Coverage
Lets take a look at how well we covered these metagenomes. First lets get coverage curves.
```{r, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=7}
# Plot curves using ggplot
ggplot(data=nonpareil.df, aes(x=effort/1000000000/2, y=coverage)) +
  geom_hline(yintercept = 1) +
  geom_hline(yintercept = 0.95, linetype=2) +
  geom_line(aes(group=SequenceID, color=CoreTemp_C)) +
  scale_color_gradient(low="blue", high="red") +
  lims(x=c(0,NA), y=c(0,1)) +
  labs(x="Sequencing effort (Gbp)", y="Estimated average coverage", color="Soil temp.\n(˚C)") +
  present_theme +
  facet_wrap(~FireClassification)
```

Now look at coverage for each sample (at full sequencing depth).

```{r, warning=FALSE, message=FALSE, fig.height=5, fig.width=7}
# Plot genome size by fire class
coverage_FC.wilcox = wilcox.test(x=filter(nonpareil.sum, FireClassification=="Reference")$C,
                              y=filter(nonpareil.sum, FireClassification=="FireAffected")$C,
                              conf.int=TRUE, conf.level=0.95)
coverage_FC.wilcox

coverage_FC.plot = ggplot(data=nonpareil.sum, aes(x=FireClassification, y=C)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill=SiteID, shape=FireClassification), size=3, width=0.25, height=0) +
  annotate("text", label="< 0.001", fontface="bold", x=1.5, y=1) +
  lims(y=c(NA, 1)) +
  scale_fill_manual(values=site.col) +
  scale_shape_manual(values=FC.shape) +
  labs(x="FireClassification", y="Estimated proportion metagenome coverage") +
  present_theme +
  theme(legend.position = "none") +
  guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=2))

# Plot genome size by temperature
coverage_temp.model.old = lme(C ~ CoreTemp_C, random = ~1|SiteID, data=nonpareil.sum)
coverage_temp.model = lmer(C ~ CoreTemp_C + (1|SiteID) + (1|Year), data=nonpareil.sum)

## Get regression for plot
coverage_temp.model.reg.df = data.frame(summary(coverage_temp.model)$coefficients) %>%
  tibble::rownames_to_column(var="factor") %>%
  mutate(p_slope = ifelse(factor == "CoreTemp_C", Pr...t.., 1),
         factor = ifelse(factor == "(Intercept)", "Intercept", factor)) %>%
  mutate(p_slope = min(p_slope)) %>%
  ungroup %>%
  select(factor, Estimate, p_slope) %>%
  tidyr::spread(key=factor, value = Estimate) %>%
  mutate(sig = ifelse(p_slope < 0.05, "< 0.05", "≥ 0.05"))
summary(coverage_temp.model)

## Plot
coverage_temp.plot = ggplot(data=nonpareil.sum, aes(x=CoreTemp_C, y=C)) +
  geom_point(aes(fill=SiteID, shape=FireClassification), size=3) +
  geom_abline(data=coverage_temp.model.reg.df, aes(intercept = Intercept, slope = CoreTemp_C), 
              linetype = 1, size=2, color="black") +
  geom_abline(data=coverage_temp.model.reg.df, aes(intercept = Intercept, slope = CoreTemp_C, linetype = sig), 
              size=1, color="white") +
  lims(y=c(NA, 1)) +
  scale_fill_manual(values=site.col) +
  scale_shape_manual(values=FC.shape) +
  scale_linetype_manual(values=c("< 0.05" = 1, "≥ 0.05" = 2)) +
  #lims(y=c(4,7)) +
  labs(x="Soil temperature (˚C)", y="Estimated proportion metagenome coverage", linetype="Regression\nslope p-value") +
  present_theme +
  guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=2),
         linetype="none")

cowplot::plot_grid(coverage_FC.plot,
                   coverage_temp.plot + theme(axis.title.y=element_blank()), 
                   rel_widths = c(0.5, 1))
```

## Diversity
Now lets take a look at the diversity.

```{r, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=7}
# Now look at diversity over fire classification
nonpareil_diversity_FC.wilcox = wilcox.test(x=filter(nonpareil.sum, FireClassification=="Reference")$diversity,
                                            y=filter(nonpareil.sum, FireClassification=="FireAffected")$diversity,
                                            conf.int=TRUE, conf.level=0.95)
nonpareil_diversity_FC.wilcox

# Plot by fire classification
nonpareil_diversity_FC.plot = ggplot(data=nonpareil.sum, aes(x=FireClassification, y=diversity)) +
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(fill=SiteID, shape=FireClassification), size=2, width=0.25) +
  annotate("text", label="< 0.001", fontface="bold", x=1.5, y=22) +
  lims(y=c(NA, 22)) +
  scale_fill_manual(values=site.col) +
  scale_shape_manual(values=FC.shape) +
  labs(x="FireClassification", y="Nonpareil diversity") +
  present_theme +
  theme(legend.position = "none") +
  guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=2))

# Plot by temperature

# Plot genome size by temperature
diversity_temp.model.old = lme(diversity ~ CoreTemp_C, random = ~1|SiteID, data=nonpareil.sum)
diversity_temp.model = lmer(diversity ~ CoreTemp_C + (1|SiteID) + (1|Year), data=nonpareil.sum)

## Get regression for plot
diversity_temp.model.reg.df = data.frame(summary(diversity_temp.model)$coefficients) %>%
  tibble::rownames_to_column(var="factor") %>%
  mutate(p_slope = ifelse(factor == "CoreTemp_C", Pr...t.., 1),
         factor = ifelse(factor == "(Intercept)", "Intercept", factor)) %>%
  mutate(p_slope = min(p_slope)) %>%
  ungroup %>%
  select(factor, Estimate, p_slope) %>%
  tidyr::spread(key=factor, value = Estimate) %>%
  mutate(sig = ifelse(p_slope < 0.05, "< 0.05", "≥ 0.05"))
summary(diversity_temp.model)

## Note I'm not running a linear model because as you'll see this is definitely not linear.
nonpareil_diversity_temp.plot = ggplot(data=nonpareil.sum, aes(x=CoreTemp_C, y=diversity)) +
  geom_point(aes(fill=SiteID, shape=FireClassification), size=2) +
  geom_abline(data=diversity_temp.model.reg.df, aes(intercept = Intercept, slope = CoreTemp_C), 
              linetype = 1, size=2, color="black") +
  geom_abline(data=diversity_temp.model.reg.df, aes(intercept = Intercept, slope = CoreTemp_C, linetype = sig), 
              size=1, color="white") +
  lims(y=c(NA, 22)) +
  scale_fill_manual(values=site.col) +
  scale_shape_manual(values=FC.shape) +
  labs(x="Soil temperature (˚C)", y="Nonpareil diversity") +
  present_theme +
  guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=2))

cowplot::plot_grid(nonpareil_diversity_FC.plot,
                   nonpareil_diversity_temp.plot + theme(axis.title.y=element_blank()), 
                   rel_widths = c(0.5, 1))
```

Lets also see how diversity changes over time.
```{r, warning=FALSE, message=FALSE, fig.height=5, fig.width=7}
# Compare across time
nonpareil_diversity_time.model.df = data.frame()
for (FC in c("FireAffected", "Reference")){
  sub_nonpareil_diversity.model = lme(diversity ~ Year, random = ~1|SiteID, data=filter(nonpareil.sum, FireClassification == FC))
  nonpareil_diversity_time.model.df = rbind(nonpareil_diversity_time.model.df,
                           data.frame(summary(sub_nonpareil_diversity.model)$tTable) %>%
                             tibble::rownames_to_column(var="factor") %>%
                             mutate(FireClassification = FC))
}
nonpareil_diversity_time.model.df


# Get regressions for plot
nonpareil_diversity_time.model.reg = nonpareil_diversity_time.model.df %>%
  mutate(p_slope = ifelse(factor == "Year", p.value, 1),
         factor = ifelse(factor == "(Intercept)", "Intercept", factor)) %>%
  group_by(FireClassification) %>%
  mutate(p_slope = min(p_slope)) %>%
  ungroup %>%
  select(FireClassification, factor, Value, p_slope) %>%
  tidyr::spread(key="factor", value="Value") %>%
  mutate(sig = ifelse(p_slope < 0.05, "< 0.05", "≥ 0.05"))

# Plot
ggplot(data=nonpareil.sum, aes(x=Year, y=diversity)) +
  geom_line(aes(group=SiteID), color="black", size=1) + 
  geom_line(aes(color=SiteID), size=0.5) + 
  geom_point(size=2, aes(fill=SiteID, shape=FireClassification)) +
  geom_abline(data=nonpareil_diversity_time.model.reg, 
              aes(intercept = Intercept, slope = Year), linetype = 1, size=2, color="black") +
  geom_abline(data=nonpareil_diversity_time.model.reg, 
              aes(intercept = Intercept, slope = Year, linetype = sig), size=1, color="white") +
  scale_shape_manual(values=FC.shape) +
  scale_fill_manual(values=site.col) +
  scale_color_manual(values=site.col) +
  scale_linetype_manual(values=c("< 0.05" = 1, "≥ 0.05" = 2)) +
  labs(x="Year", y="Nonpareil diversity", linetype="LME p-value") +
  present_theme +
  theme(legend.direction = "vertical") +
  facet_wrap(~FireClassification) +
  guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=2),
         linetype=guide_legend(override.aes=list(color="black")))

```

# Estimated genome size

We used microbecensus to estimate average genome size.

```{r, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=7}
# Import microbecensus data
microbe_census.df = read.table("/Users/sambarnett/Documents/Shade_lab/Centralia_project/Metagenomics/Data/microbecensus_comb_out.txt",
                               header=TRUE, sep="\t") %>%
  mutate(average_genome_size_Mbp = average_genome_size/1000000) %>%
  arrange(-average_genome_size) %>%
  mutate(Gsize_rank = row_number()) %>%
  left_join(sample.meta, by = "SequenceID") %>%
  left_join(readstats.df, by = "SequenceID")

# Plot genome size by fire class
Gsize_FC.wilcox = wilcox.test(x=filter(microbe_census.df, FireClassification=="Reference")$average_genome_size_Mbp,
                              y=filter(microbe_census.df, FireClassification=="FireAffected")$average_genome_size_Mbp,
                              conf.int=TRUE, conf.level=0.95)
Gsize_FC.wilcox

Gsize_FC.plot = ggplot(data=microbe_census.df, aes(x=FireClassification, y=average_genome_size_Mbp)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill=SiteID, shape=FireClassification), size=3, width=0.25, height=0) +
  annotate("text", label="< 0.001", fontface="bold", x=1.5, y=6.75) +
  lims(y=c(NA, 6.75)) +
  scale_fill_manual(values=site.col) +
  scale_shape_manual(values=FC.shape) +
  labs(x="FireClassification", y="Estimated average genome size (Mbp)") +
  present_theme +
  theme(legend.position = "none") +
  guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=2))

# Plot genome size by temperature
Gsize_temp.model.old = lme(average_genome_size_Mbp ~ CoreTemp_C, random = ~1|SiteID, data=microbe_census.df)
Gsize_temp.model = lmer(average_genome_size_Mbp ~ CoreTemp_C + (1|SiteID) + (1|Year), data=microbe_census.df)

## Get regression for plot
Gsize_temp.model.reg.df = data.frame(summary(Gsize_temp.model)$coefficients) %>%
  tibble::rownames_to_column(var="factor") %>%
  mutate(p_slope = ifelse(factor == "CoreTemp_C", Pr...t.., 1),
         factor = ifelse(factor == "(Intercept)", "Intercept", factor)) %>%
  mutate(p_slope = min(p_slope)) %>%
  ungroup %>%
  select(factor, Estimate, p_slope) %>%
  tidyr::spread(key=factor, value = Estimate) %>%
  mutate(sig = ifelse(p_slope < 0.05, "< 0.05", "≥ 0.05"))
summary(Gsize_temp.model)

## Plot
Gsize_temp.plot = ggplot(data=microbe_census.df, aes(x=CoreTemp_C, y=average_genome_size_Mbp)) +
  geom_point(aes(fill=SiteID, shape=FireClassification), size=3) +
  geom_abline(data=Gsize_temp.model.reg.df, aes(intercept = Intercept, slope = CoreTemp_C), 
              linetype = 1, size=2, color="black") +
  geom_abline(data=Gsize_temp.model.reg.df, aes(intercept = Intercept, slope = CoreTemp_C, linetype = sig), 
              size=1, color="white") +
  lims(y=c(NA, 6.75)) +
  scale_fill_manual(values=site.col) +
  scale_shape_manual(values=FC.shape) +
  scale_linetype_manual(values=c("< 0.05" = 1, "≥ 0.05" = 2)) +
  #lims(y=c(4,7)) +
  labs(x="Soil temperature (˚C)", y="Estimated average genome size (Mbp)", linetype="Regression\nslope p-value") +
  present_theme +
  guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=2),
         linetype="none")

cowplot::plot_grid(Gsize_FC.plot,
                   Gsize_temp.plot + theme(axis.title.y=element_blank()), 
                   rel_widths = c(0.5, 1))
```

Now lets look at this over time like before

```{r, warning=FALSE, message=FALSE, fig.height=5, fig.width=7}
# Compare across time
Gsize_time.model.df = data.frame()
for (FC in c("FireAffected", "Reference")){
  sub_Gsize.model = lme(average_genome_size_Mbp ~ Year, random = ~1|SiteID, data=filter(microbe_census.df, FireClassification == FC))
  Gsize_time.model.df = rbind(Gsize_time.model.df,
                           data.frame(summary(sub_Gsize.model)$tTable) %>%
                             tibble::rownames_to_column(var="factor") %>%
                             mutate(FireClassification = FC))
}
Gsize_time.model.df

## Get regression for plot
Gsize_time.model.reg = Gsize_time.model.df %>%
  mutate(p_slope = ifelse(factor == "Year", p.value, 1),
         factor = ifelse(factor == "(Intercept)", "Intercept", factor)) %>%
  group_by(FireClassification) %>%
  mutate(p_slope = min(p_slope)) %>%
  ungroup %>%
  select(FireClassification, factor, Value, p_slope) %>%
  tidyr::spread(key="factor", value="Value") %>%
  mutate(sig = ifelse(p_slope < 0.05, "< 0.05", "≥ 0.05"))

## Plot
ggplot(data=microbe_census.df, aes(x=Year, y=average_genome_size_Mbp)) +
  geom_line(aes(group=SiteID), color="black", size=1) + 
  geom_line(aes(color=SiteID), size=0.5) + 
  geom_point(size=2, aes(fill=SiteID, shape=FireClassification)) +
  geom_abline(data=Gsize_time.model.reg, 
              aes(intercept = Intercept, slope = Year), linetype = 1, size=2, color="black") +
  geom_abline(data=Gsize_time.model.reg, 
              aes(intercept = Intercept, slope = Year, linetype = sig), size=1, color="white") +
  scale_shape_manual(values=FC.shape) +
  scale_fill_manual(values=site.col) +
  scale_color_manual(values=site.col) +
  scale_linetype_manual(values=c("< 0.05" = 1, "≥ 0.05" = 2)) +
  labs(x="Year", y="Estimated average genome size (Mbp)", linetype="LME p-value") +
  present_theme +
  theme(legend.direction = "vertical") +
  facet_wrap(~FireClassification) +
  guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=2),
         linetype=guide_legend(override.aes=list(color="black")))

```

## Coverage, diversity and genome size plots for publication
Lets make a nice plot with both results for publication

```{r, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# Coverage plots
coverage_FC.plot = ggplot(data=nonpareil.sum, aes(x=FireClassification, y=C)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill=SiteID, shape=FireClassification), size=2, width=0.25, height=0) +
  annotate("text", label="< 0.001", x=1.5, y=1, size=6*5/14) +
  lims(y=c(NA, 1)) +
  scale_fill_manual(values=site.col) +
  scale_shape_manual(values=FC.shape) +
  labs(x="FireClassification", y="Estimated proportion\nmetagenome coverage") +
  publication_theme +
  theme(legend.position = "none") +
  guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=2))

coverage_temp.plot = ggplot(data=nonpareil.sum, aes(x=CoreTemp_C, y=C)) +
  geom_point(aes(fill=SiteID, shape=FireClassification), size=2) +
  geom_abline(data=filter(coverage_temp.model.reg.df, p_slope < 0.05), 
              aes(intercept = Intercept, slope = CoreTemp_C), 
              linetype = 2, size=1, color="black") +
  lims(y=c(NA, 1)) +
  scale_fill_manual(values=site.col) +
  scale_shape_manual(values=FC.shape) +
  scale_linetype_manual(values=c("< 0.05" = 1, "≥ 0.05" = 2)) +
  #lims(y=c(4,7)) +
  labs(x="Soil temperature (˚C)", y="Estimated proportion\nmetagenome coverage", linetype="Regression\nslope p-value") +
  publication_theme +
  guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=2),
         linetype="none")

coverage.plots = cowplot::plot_grid(coverage_FC.plot,
                                    coverage_temp.plot + theme(axis.title.y=element_blank(), legend.position = "none"), 
                                    rel_widths = c(0.7, 1))

# Diversity plots
nonpareil_diversity_FC.plot = ggplot(data=nonpareil.sum, aes(x=FireClassification, y=diversity)) +
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(fill=SiteID, shape=FireClassification), size=2, width=0.25) +
  annotate("text", label="< 0.001", x=1.5, y=22, size=6*5/14) +
  lims(y=c(NA, 22)) +
  scale_fill_manual(values=site.col) +
  scale_shape_manual(values=FC.shape) +
  labs(x="FireClassification", y="Nonpareil diversity") +
  publication_theme +
  theme(legend.position = "none") +
  guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=2))

nonpareil_diversity_temp.plot = ggplot(data=nonpareil.sum, aes(x=CoreTemp_C, y=diversity)) +
  geom_point(aes(fill=SiteID, shape=FireClassification), size=2) +
  geom_abline(data=filter(diversity_temp.model.reg.df, p_slope < 0.05), 
              aes(intercept = Intercept, slope = CoreTemp_C), 
              linetype = 2, size=1, color="black") +
  lims(y=c(NA, 22)) +
  scale_fill_manual(values=site.col) +
  scale_shape_manual(values=FC.shape) +
  labs(x="Soil temperature (˚C)", y="Nonpareil diversity") +
  publication_theme +
  guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=2))

nonpareil_diversity.plots = cowplot::plot_grid(nonpareil_diversity_FC.plot,
                                               nonpareil_diversity_temp.plot + theme(axis.title.y=element_blank(), legend.position = "none"), 
                                               rel_widths = c(0.7, 1))

# Genome size plots
Gsize_FC.plot = ggplot(data=microbe_census.df, aes(x=FireClassification, y=average_genome_size_Mbp)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill=SiteID, shape=FireClassification), size=2, width=0.25, height=0) +
  annotate("text", label="< 0.001", x=1.5, y=6.75, size=6*5/14) +
  lims(y=c(NA, 6.75)) +
  scale_fill_manual(values=site.col) +
  scale_shape_manual(values=FC.shape) +
  labs(x="FireClassification", y="Estimated average\ngenome size (Mbp)") +
  publication_theme +
  theme(legend.position = "none") +
  guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=2))

Gsize_temp.plot = ggplot(data=microbe_census.df, aes(x=CoreTemp_C, y=average_genome_size_Mbp)) +
  geom_point(aes(fill=SiteID, shape=FireClassification), size=2) +
  geom_abline(data=filter(Gsize_temp.model.reg.df, p_slope < 0.05), 
              aes(intercept = Intercept, slope = CoreTemp_C), 
              linetype = 2, size=1, color="black") +
  lims(y=c(NA, 6.75)) +
  scale_fill_manual(values=site.col) +
  scale_shape_manual(values=FC.shape) +
  scale_linetype_manual(values=c("< 0.05" = 1, "≥ 0.05" = 2)) +
  #lims(y=c(4,7)) +
  labs(x="Soil temperature (˚C)", y="Estimated average\ngenome size (Mbp)", linetype="Regression\nslope p-value") +
  publication_theme +
  guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=2),
         linetype="none")

Gsize.plots = cowplot::plot_grid(Gsize_FC.plot,
                                 Gsize_temp.plot + theme(axis.title.y=element_blank(), legend.position = "none"), 
                                 rel_widths = c(0.7, 1))

# Plot all together
read_sum.plots = cowplot::plot_grid(coverage.plots,
                                    nonpareil_diversity.plots,
                                    Gsize.plots, ncol=1, labels = c("A", "B", "C"),
                                    label_size = 8)
read_sum.plots = cowplot::plot_grid(read_sum.plots, 
                                    g_legend(Gsize_temp.plot + guides(fill=guide_legend(override.aes=list(shape=site.shape), ncol=1))), 
                                    nrow=1, rel_widths = c(1,0.3))


read_sum.plots

ggsave(read_sum.plots, file="/Users/sambarnett/Documents/Shade_lab/Centralia_project/Metagenomics/Manuscript/Revision_1/Figures/Supplemental/FigS1.tiff",
       device="tiff", width=5, height=5, units="in", bg="white")
```


# Taxonomy

## Read based taxonomy
We estimated read based taxonomy using braken Lets take a look at the broad taxonomic breakdown of these metagenomes.
```{r, warning=FALSE, message=FALSE, fig.height=5, fig.width=10}
# Read in taxonomy IDs and their taxonomy breakdown used by braken. We will use these to link the taxonomy IDs from the braken output to taxonomy.
Main_taxonomy.df = read.table("/Users/sambarnett/Documents/Shade_lab/Centralia_project/Metagenomics/Data/Annotations/main_taxonomic_ranks.txt",
                              header=TRUE, sep="\t", quote="", comment.char="") %>%
  tidyr::spread(key=Ancestor_level, value=Ancestor) %>%
  rename(Domain = D,
         Phylum = P,
         Class = C,
         Order = O,
         Family = F,
         Genus = G,
         Species = S)

# Read in the braken results and map taxonomies and summarize read percentages per class and only indicate those with over 1% of the reads.

bracken.df = read.table("/Users/sambarnett/Documents/Shade_lab/Centralia_project/Metagenomics/Data/Read_taxonomy.txt",
                               header=TRUE, sep="\t", quote="", comment.char="") %>%
  rename(TaxID = taxonomy_id) %>%
  left_join(Main_taxonomy.df, by="TaxID") %>%
  filter(Domain == "Bacteria") %>%
  group_by(SequenceID) %>%
  mutate(total_classified_reads = sum(fraction_total_reads)) %>%
  ungroup %>%
  mutate(percent_class_reads = fraction_total_reads/total_classified_reads*100) %>%
  group_by(Domain, Class) %>%
  mutate(max_perc_reads = max(percent_class_reads)) %>%
  ungroup %>%
  mutate(Class = ifelse(max_perc_reads < 1, "Less than 1%", Class)) %>%
  group_by(SequenceID, Class) %>%
  summarize(percent_class_reads = sum(percent_class_reads)) %>%
  ungroup %>%
  left_join(sample.meta, by = "SequenceID") %>%
  mutate(SequenceID = factor(SequenceID, levels = arrange(sample.meta, CoreTemp_C)$SequenceID))
bracken.df$Class = factor(bracken.df$Class,
                          levels = c(sort(unique(filter(bracken.df, Class != "Less than 1%")$Class)), "Less than 1%"))

class.col = c(paultol_colors(length(levels(bracken.df$Class))-1), "#777777")
names(class.col) = levels(bracken.df$Class)

# Plot (with temperature for fun)
ggplot(data=bracken.df, aes(x=as.factor(Year), y=percent_class_reads)) +
  geom_bar(stat="identity", aes(fill=Class), color="black") +
  geom_line(data=sample.meta, aes(y=CoreTemp_C/0.5, group=1), size=1, color="black") +
  geom_line(data=sample.meta, aes(y=CoreTemp_C/0.5, group=1), size=0.5, color="green") +
  scale_fill_manual(values = class.col) +
  scale_y_continuous(name = "Percent of Bacterial reads (%)", limits = c(-4, 101), 
                    sec.axis = sec_axis(~.*0.5, name="Soil temperature (˚C)")) +
  labs(x="Year", fill="Bacterial class") +
  present_theme +
  theme(axis.text.x = element_text(angle=90)) +
  facet_wrap(~FireClassification*SiteID, nrow=2)

```

## Contig taxonomy

From the assembled contigs we used kraken to taxonomically assign. Lets see the taxonomic breakdown of these assemblies.
```{r, warning=FALSE, message=FALSE, fig.height=5, fig.width=7}
# Read in the taxonomy from the IDs
Main_taxonomy.df = read.table("/Users/sambarnett/Documents/Shade_lab/Centralia_project/Metagenomics/Data/Annotations/main_taxonomic_ranks.txt",
                              header=TRUE, sep="\t", quote="", comment.char="") %>%
  tidyr::spread(key=Ancestor_level, value=Ancestor) %>%
  rename(Domain = D,
         Phylum = P,
         Class = C,
         Order = O,
         Family = F,
         Genus = G,
         Species = S)

# Read in the contig taxonomic assignments and match with their IDs
contig_tax.df = read.table("/Users/sambarnett/Documents/Shade_lab/Centralia_project/Metagenomics/Data/Annotations/Contig_taxonomy.txt",
                               header=TRUE, sep="\t", quote="", comment.char="") %>%
  rename(TaxID = TaxonomyID) %>%
  left_join(Main_taxonomy.df, by="TaxID") %>%
  mutate(Taxa = ifelse(is.na(Domain) | Domain == "Unclassified", "Unclassified",
                       ifelse(is.na(Phylum) | Phylum == "Unclassified", paste("Unclassified", Domain),
                              paste(Domain, Phylum, sep = "; "))))

# Summarize contig percentages per phyla and only indicate those with over 1% of the contigs
classified_contig.sum = contig_tax.df %>%
  group_by(SequenceID) %>%
  mutate(total_contigs = n()) %>%
  ungroup %>%
  filter(!grepl("Unclassified", Taxa)) %>%
  group_by(SequenceID, total_contigs) %>%
  summarize(classified_contigs = n()) %>%
  ungroup %>%
  mutate(percent_classified_contigs = classified_contigs/total_contigs*100) %>%
  left_join(sample.meta, by = "SequenceID")

contig_tax.sum = contig_tax.df %>%
  filter(!grepl("Unclassified", Taxa)) %>%
  group_by(SequenceID) %>%
  mutate(classified_contigs = n()) %>%
  ungroup %>%
  group_by(SequenceID, classified_contigs, Taxa) %>%
  summarize(n_contigs = n()) %>%
  ungroup %>%
  mutate(percent_contigs = n_contigs/classified_contigs*100) %>%
  group_by(Taxa) %>%
  mutate(max_percent_contigs = max(percent_contigs)) %>%
  ungroup %>%
  mutate(Taxa = ifelse(max_percent_contigs < 1, "Less than 1%", Taxa)) %>%
  group_by(SequenceID, Taxa) %>%
  summarize(percent_contigs = sum(percent_contigs)) %>%
  ungroup %>%
  left_join(sample.meta, by = "SequenceID")

contig_tax.sum$Taxa = factor(contig_tax.sum$Taxa,
                          levels = c(sort(unique(filter(contig_tax.sum, Taxa != "Less than 1%")$Taxa)), "Less than 1%"))

taxa.col = c(paultol_colors(length(levels(contig_tax.sum$Taxa))-1), "#777777")
names(taxa.col) = levels(contig_tax.sum$Taxa)

# Plot (with temperature for fun)
contig_tax.plot = ggplot(data=contig_tax.sum, aes(x=as.factor(Year), y=percent_contigs)) +
  geom_bar(stat="identity", aes(fill=Taxa), color="black", size=0.3) +
  geom_line(data=sample.meta, aes(y=CoreTemp_C/0.5, group=1), size=1, color="black") +
  geom_line(data=sample.meta, aes(y=CoreTemp_C/0.5, group=1), size=0.5, color="white") +
  #geom_text(data=classified_contig.sum, aes(label=round(percent_classified_contigs, digits = 1)), 
  #          y=102, size=5*5/14, angle=90, hjust=0) +
  scale_fill_manual(values = taxa.col) +
  scale_y_continuous(name = "Percent of phylum-classified contigs (%)", limits = c(-1, 101), 
                    sec.axis = sec_axis(~.*0.5, name="Soil temperature (˚C)")) +
  labs(x="Year", fill="Domain; Phylum") +
  publication_theme +
  theme(axis.text.x = element_text(angle=90)) +
  facet_wrap(~FireClassification*SiteID, nrow=2)
contig_tax.plot

#ggsave(contig_tax.plot, file="/Users/sambarnett/Documents/Shade_lab/Centralia_project/Metagenomics/Resource announcement/Fig1.tiff",
#       device="tiff", width=7, height=5, units="in", bg="white")

```

# Sequencing stats for resource announcement

Here are some sequencing stats used for the resource announcement.

```{r, warning=FALSE, message=FALSE}
MRA.df = sample.meta %>%
  mutate(Collection_date = paste(Day, "Oct", Year)) %>%
  select(SequenceID, SiteID, Year, Collection_date, CoreTemp_C, FireClassification) %>%
  left_join(read.table("/Users/sambarnett/Documents/Shade_lab/Centralia_project/Metagenomics/Data/Contig_summaries.txt",
                       header=TRUE, sep="\t", quote="", comment.char="")) %>%
  left_join(readstats.df) %>%
  arrange(SiteID, Year) %>%
  select(-Year)
MRA.df
#write.table(MRA.df, file="/Users/sambarnett/Documents/Shade_lab/Centralia_project/Metagenomics/Resource announcement/Table1.txt",
#            sep="\t", quote=FALSE, row.names = FALSE)
```

# Centralia Map for revision
Now lets make a map of Centralia and all our sites for manuscript revision.

```{r}
library(ggmap)
library(grid)
library(sf)

# Get site coordinates. These are found in a table with longitude (long) and latitude (lat) for each site.
site.coords = read_xlsx("/Users/sambarnett/Documents/Shade_lab/Centralia_project/Centralia_soil_metadata.xlsx", sheet = "Site_metadata")  %>%
  mutate(longitude = -1*as.numeric(measurements::conv_unit(longitude, from = 'deg_dec_min', to = 'dec_deg')),
         latitude = as.numeric(measurements::conv_unit(latitude, from = 'deg_dec_min', to = 'dec_deg'))) %>%
  filter(SiteID %in% used_sites)

## Getting the center point of the cite locations.
center = c(lon=mean(c(max(site.coords$longitude), min(site.coords$longitude))), 
           lat=mean(c(max(site.coords$latitude), min(site.coords$latitude))))

## Get the 4 edges of the region around the above center point.
Nlat = geosphere::destPoint(center, b=0, d=150)[2]
Slat = geosphere::destPoint(center, b=180, d=150)[2]
Elong = geosphere::destPoint(center, b=90, d=250)[1]
Wlong = geosphere::destPoint(center, b=270, d=250)[1]

## Making a polygon (rectangle) of these edges.
region.poly = data.frame(corner = c("NE", "SE", "SW", "NW"), 
                         lat = c(Nlat, Slat, Slat, Nlat), 
                         long = c(Elong, Elong, Wlong, Wlong),
                         group=3)

# Front coordinates
front_coord.df = data.frame(front = c(1, 2),
                            lon_start = -76.332739, lat_start = 40.801547,
                            lon_end = c(-76.345905, -76.344572),
                            lat_end = c(40.801056, 40.796094)) %>%
  mutate(m = (lat_end-lat_start)/(lon_end-lon_start)) %>%
  mutate(b = lat_start-(m*lon_start))
```

Make the Pennsylvania map.

```{r}
## Make state map for inset. The field sites were in PA, USA so we want to easily show where in the state the regional map (above polygon) is located
pa.coords = map_data("state", "pennsylvania")

pa.region.map = ggplot(data = pa.coords, aes(x=long, y = lat, group = group)) + 
  geom_polygon(fill="grey", color="black", size=0.25) + 
  #geom_polygon(data=region.poly, fill="red") +
  annotate("point", x=center["lon"], y=center["lat"], fill="red", color="black", shape=21, size=1) +
  theme_bw() +
  publication_theme +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.text=element_blank(),
        axis.title=element_blank(),
        axis.ticks=element_blank(),
        plot.background=element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank()) +
  coord_fixed(1.3)
pa.region.map
```

Make a USA map.

```{r}
## Make state map for inset. The field sites were in PA, USA so we want to easily show where in the state the regional map (above polygon) is located
usa.coords = map_data("usa")

pa_usa.region.map = ggplot(data = usa.coords, aes(x=long, y = lat, group = group)) + 
  geom_polygon(fill="white", color="black", size=0.25) + 
  geom_polygon(data=pa.coords, fill="grey", color="black", size=0.25) + 
  #geom_polygon(data=region.poly, fill="red") +
  #annotate("point", x=center["lon"], y=center["lat"], fill="orange", color="black", shape=21, size=1) +
  theme_bw() +
  publication_theme +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.text=element_blank(),
        axis.title=element_blank(),
        axis.ticks=element_blank(),
        plot.background=element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank()) +
  coord_fixed(1.3)
pa_usa.region.map
```

Now make the Centralia map and put it all together

```{r, fig.width=3.5, fig.height=5}
## Make map of the specific region using the stamen terrain map.
sbbox <- make_bbox(lon = c(Elong, Wlong), lat = c(Slat, Nlat), f = .1)
region.map = get_googlemap(center = c(lon=-76.343139, lat=40.800804),
                           maptype="satellite", zoom=16, scale=2) %>% 
  ggmap()


site.coords$lon = site.coords$longitude
site.coords$lat = site.coords$latitude

xmax = max(layer_scales(region.map)$x$range$range)
region.points.map = region.map +
  geom_segment(data=filter(front_coord.df, front == 1), x=xmax,
               aes(y=(m*xmax + b), xend=lon_end, yend=lat_end),
               arrow = arrow(length = unit(0.05, "npc")), color="white") +
  geom_point(data=filter(site.coords, SiteID %in% used_sites), 
             aes(x=lon, y=lat, fill=SiteID, shape=FireClassification), size=2, color="white") +
  scale_shape_manual(values=FC.shape) +
  scale_fill_manual(values=site.col) +
  labs(x="Longitude", y="Latitude", fill="SiteID", shape="Fire classification") +
  lims(x=sbbox[c(1,3)], y=sbbox[c(2,4)]) +
  theme_bw() +
  publication_theme +
  theme(legend.position = "bottom") +
  guides(fill=guide_legend(override.aes=list(shape=site.shape, color="black"), title.position="top", title.hjust=0.5, ncol=3),
         shape=guide_legend(override.aes=list(color="black"), title.position="top", title.hjust=0.5, ncol=1))

map.leg = g_legend(region.points.map)

# Combine the maps together
USA_PA.map = cowplot::plot_grid(pa_usa.region.map, pa.region.map, nrow=1)

Centralia.map = cowplot::plot_grid(USA_PA.map, 
                                   region.points.map + theme(legend.position="none"),
                                   map.leg, ncol=1, rel_heights = c(0.5,1,0.5))
Centralia.map
```


# New Figure S1
Combine original figure S1 with map
```{r, fig.width=7, fig.height=5}
# Plot all together
read_sum.plots = cowplot::plot_grid(coverage.plots,
                                    nonpareil_diversity.plots,
                                    Gsize.plots, ncol=1, labels = c("B", "C", "D"),
                                    label_size = 8)
read_sum.plots = cowplot::plot_grid(Centralia.map, read_sum.plots,
                                    nrow=1, labels = c("A", ""),
                                    label_size = 8)
read_sum.plots

ggsave(read_sum.plots, file="/Users/sambarnett/Documents/Shade_lab/Centralia_project/Metagenomics/Manuscript/Revision_1/Figures/Supplemental/FigS1.tiff",
       device="tiff", width=7, height=5, units="in", bg="white")
```




# Session info

```{r}
sessionInfo()
```

